Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-03-17 paper.qmd
Added function to get 2D projection data

@@

two_curvy_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.6, 0.6)) + 
  ylim(c(-0.6, 0.6)) +
  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
  theme(
    legend.position = "none"
  )

@@

two_curvy_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("a3", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

@@

projection_scaled <- projection * 2.5

projected <- as.matrix(scaled_mnist_data) %*% projection_scaled

projected_df <- projected |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  #dplyr::mutate(type = df_exe$type) |>
  dplyr::mutate(ID = dplyr::row_number()) |>
  mutate(cluster = tsne_minst_scaled_temp$cluster)

projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled

model_df <- dplyr::left_join(
  distance_df_small_edges_mnist |> select(-distance), 
  projected_model_df, 
  by = c("from" = "ID"))

axes_obj <- gen_axes(  
  proj = projection,
  limits = 1,
  axis_pos_x = -0.9,
  axis_pos_y = -0.9,
  axis_labels = names(scaled_mnist_data),
  threshold = 0.06)

@@

projection_scaled <- projection * 2

projected <- as.matrix(scaled_mnist_data) %*% projection_scaled

projected_df <- projected |>
  tibble::as_tibble(.name_repair = "unique") |>
  dplyr::rename(c("proj1" = "...1",
                  "proj2" = "...2")) |>
  #dplyr::mutate(type = df_exe$type) |>
  dplyr::mutate(ID = dplyr::row_number()) |>
  mutate(cluster = tsne_minst_scaled_temp$cluster)

projected_model <- as.matrix(scaled_mnist_data_model) %*% projection_scaled

model_df <- dplyr::left_join(
  distance_df_small_edges_mnist |> select(-distance), 
  projected_model_df, 
  by = c("from" = "ID"))

axes_obj <- gen_axes(  
  proj = projection,
  limits = 1,
  axis_pos_x = -0.83,
  axis_pos_y = -0.83,
  axis_labels = names(scaled_mnist_data),
  threshold = 0.06)

>>>>

```{r}
#| label: plot-projection

get_projection <- function(projection, proj_scale, scaled_data, 
                           scaled_data_model, distance_df_small_edges,
                           axis_param) {

  projection_scaled <- projection * proj_scale
  projected <- as.matrix(scaled_data) %*% projection_scaled

  projected_model <- as.matrix(scaled_data_model) %*% projection_scaled

  limits <- axis_param$limits
  axis_pos_x <- axis_param$axis_pos_x
  axis_pos_y <- axis_param$axis_pos_y
  threshold <- axis_param$threshold

  axes_obj <- gen_axes(  
    proj = projection,
    limits = limits,
    axis_pos_x = axis_pos_x,
    axis_pos_y = axis_pos_y,
    axis_labels = names(scaled_data),
    threshold = threshold)

  return(list(projected_df = projected_df, 
              model_df = model_df, 
              axes = axes, 
              circle = circle))

@@

two_curvy_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.6, 0.6)) + 
  ylim(c(-0.6, 0.6)) +
  interior_annotation("a", c(0.08, 0.93), cex = 1.5) +
  theme(
    legend.position = "none"
  )

@@

two_curvy_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.2,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 5) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.1, 0.1)) + 
  ylim(c(-0.1, 0.1)) +
  interior_annotation("a3", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

@@

proj_obj1 <- get_projection(projection = projection, 
               proj_scale = 2.5, 
               scaled_data = scaled_mnist_data, 
               scaled_data_model = scaled_mnist_data_model, 
               distance_df_small_edges = distance_df_small_edges_mnist, 
               axis_param = list(limits = 1, 
                                 axis_pos_x = -0.9, 
                                 axis_pos_y = -0.9, 
                                 threshold = 0.06))

projected_df <- proj_obj1$projected_df |>
  mutate(cluster = tsne_minst_scaled_temp$cluster)

model_df <- proj_obj1$model_df
axes <- proj_obj1$axes
circle <- proj_obj1$circle

@@

proj_obj2 <- get_projection(projection = projection, 
               proj_scale = 2, 
               scaled_data = scaled_mnist_data, 
               scaled_data_model = scaled_mnist_data_model, 
               distance_df_small_edges = distance_df_small_edges_mnist, 
               axis_param = list(limits = 1, 
                                 axis_pos_x = -0.83, 
                                 axis_pos_y = -0.83, 
                                 threshold = 0.06))

projected_df <- proj_obj2$projected_df |>
  mutate(cluster = tsne_minst_scaled_temp$cluster)

model_df <- proj_obj2$model_df
axes <- proj_obj2$axes
circle <- proj_obj2$circle
